Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-29 paper/main.tex
restructure appendix to fit plots template

%###################################################################################################
\section{Acknowledgements}
\subsection{Author Contributions}
\textbf{Conceptualization:} Jonas Pleyer\\
\textbf{Investigation:} Jonas Pleyer\\
\textbf{Methodology} Jonas Pleyer\\
\textbf{Software:} Jonas Pleyer\\
\textbf{Supervision:} Christian Fleck\\
\textbf{Writing – original draft:} Jonas Pleyer\\
\textbf{Writing – review \& editing:} Christian Fleck \& Jonas Pleyer

\subsection{Declaration of AI Use}
The authors acknowledge the use of Gemini (Google) for assistance with language editing and
proofreading.
In particular, no generative aspect was used and each suggested change was carefully reviewed and
utilized only as additional feedback source.
All contents of the paper are written and audited by the authors which is why they take full
responsibility for the contents presented in this publication.

\section{Data and Code Availability}
\label{section:data-code-availability}
All data and code which has been used in this study is available at the github repository
\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
The package which bundles the developed funtionality is published at
\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
dependency.\\
The time series of microscopic images which has been studied in section~\ref{section:application-2}
was extracted from individual frames of a video~\cite{Trude1982} that
showcases the exponential growth of bacteria.

% \bibliographystyle{IEEEtran}
\bibliography{references}

\newpage
\appendix
%###################################################################################################
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{S\arabic{subsection}}
\section{Supplementary Material}

%###################################################################################################
\subsection{Alternative Interaction Potential - Morse Potential}
\label{sec:supplement-morse-potential}
The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
potential.
It was originally designed to describe the interaction of diatomic molecules but has since found use
in other areas~\cite{Breitwieser2021}.
It contains a repulsive and attractive part and is bound from above which makes it particularly
suitable for numerical simulations since unbound potentials can lead to large numerical deviations
thus compromising the prediction of the model.
The potential is of the form
\begin{equation}
    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
\end{equation}
and has 3 parameters.
The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
As done with our approach towards the Mie potential, we are free to assign a spatial cutoff $\xi$
to the range of the interaction potential.

\subsection{Further Details of the Extraction Algorithm}
\label{sec:supplement-extraction-algorithm}

@@

%###################################################################################################
\subsection{Analysis \& Usage of heterogeneous Parameter Value (TODO)}
In order to reuse this model, we also provide the option to assign growth rates by sampling from a
normal distribution with user-defined parameters.

%###################################################################################################
\subsection{Application Data \& Preparation (TODO)}
\label{sec:supplement-data-details}

@@

%###################################################################################################
\subsection{Parameter Estimation Results (TODO)}
\label{sec:supplement-parameter-profiles}

@@

%###################################################################################################
\subsection{Implementation \& Performance (TODO)}
\label{section:supplement-performance}

@@

>>>>

\newpage
%###################################################################################################
% \addcontentsline{toc}{chapter}{Supporting Information}
% \section*{Supplementary Material}
% \renewcommand{\thesection}{S\arabic{section}}
% \renewcommand{\thesubsection}{S\arabic{subsection}}

\section*{Supporting Information}

%###################################################################################################
\paragraph*{Alternative Interaction Potential - Morse Potential}
\label{sec:supplement-morse-potential}
The morse potential~\cite{Morse1929} is a common example of numerically inexpensive interaction
potential.
It was originally designed to describe the interaction of diatomic molecules but has since found use
in other areas~\cite{Breitwieser2021}.
It contains a repulsive and attractive part and is bound from above which makes it particularly
suitable for numerical simulations since unbound potentials can lead to large numerical deviations
thus compromising the prediction of the model.
The potential is of the form
\begin{equation}
    V(r) = V_0\left(1-e^{-\lambda(r-(R_1+R_2))}\right)^2
\end{equation}
and has 3 parameters.
The overall potential strength $V_0$, the radii of the particles $R_1,R_2$ and the potential
stiffness $\lambda$ which controls the rate of change as one travels along the radial axis.
As done with our approach towards the Mie potential, we are free to assign a spatial cutoff $\xi$
to the range of the interaction potential.

\paragraph*{Further Details of the Extraction Algorithm}
\label{sec:supplement-extraction-algorithm}

@@

%###################################################################################################
\paragraph*{Analysis \& Usage of heterogeneous Parameter Value (TODO)}
In order to reuse this model, we also provide the option to assign growth rates by sampling from a
normal distribution with user-defined parameters.

%###################################################################################################
\paragraph*{Application Data \& Preparation (TODO)}
\label{sec:supplement-data-details}

@@

%###################################################################################################
\paragraph*{Parameter Estimation Results (TODO)}
\label{sec:supplement-parameter-profiles}

@@

%###################################################################################################
\paragraph*{Implementation \& Performance (TODO)}
\label{section:supplement-performance}

@@

%###################################################################################################
\section*{Acknowledgements}
\paragraph{Author Contributions}
\textbf{Conceptualization:} Jonas Pleyer\\
\textbf{Investigation:} Jonas Pleyer\\
\textbf{Methodology} Jonas Pleyer\\
\textbf{Software:} Jonas Pleyer\\
\textbf{Supervision:} Christian Fleck\\
\textbf{Writing – original draft:} Jonas Pleyer\\
\textbf{Writing – review \& editing:} Christian Fleck \& Jonas Pleyer

\paragraph*{Declaration of AI Use}
The authors acknowledge the use of Gemini (Google) for assistance with language editing and
proofreading.
In particular, no generative aspect was used and each suggested change was carefully reviewed and
utilized only as additional feedback source.
All contents of the paper are written and audited by the authors which is why they take full
responsibility for the contents presented in this publication.

\section*{Data and Code Availability}
\label{section:data-code-availability}
All data and code which has been used in this study is available at the github repository
\url{https://github.com/jonaspleyer/cr\_mech\_coli}.
It includes detailed descriptions of how to obtain the used data and bring it into a usable format.
The package which bundles the developed funtionality is published at
\url{https://pypi.org/project/cr-mech-coli/} from which it is readily available as a python
dependency.\\
The time series of microscopic images which has been studied in section~\ref{section:application-2}
was extracted from individual frames of a video~\cite{Trude1982} that
showcases the exponential growth of bacteria.

\nolinenumbers
\newpage
\bibliography{references}
